Transforming Trash as an Artistic Act

### 2015-12-29 chapter3_reduced.tex
thesis format
Former-commit-id: 4c193566fa7da80661a70ce4067ea75d9e65ec88
thesis format
Former-commit-id: 45314951bf19c52d30f0c36d2e22e1c79e2717e1

\chapter{Trash (in) Art}

\epigraph{One day, in a rubbish heap, I found an old bicycle seat lying beside a rusted handlebar, and my mind instantly linked them together. I assembled these two objects, which everyone then recognized as a bull’s head. The metamorphosis was accomplished, and I wish another metamorphosis would occur in the reverse sense. If my bull’s head were thrown in a junk heap, perhaps one day some boy would say, \quotes{Here’s something that would make a good handlebar for my bicycle!}}{\hfill ---Pablo Picasso, \textit{Trashformations}, 1998}

In this chapter, questioned the purpose and meaning of using trash as a medium in the artworks. The reasons and motivations of artists who use trash are analyzed. The artist uses discarded items in their work is analyzed in the light of arguments discussed in the previous chapter. Questioned that why artists select to use society's debris rather than fresh items. Researched the place of trash in the world of art by looking at different methods and approaches. Started from masterpieces and founding works. Continued with recent examples of artworks.

@@

Firstly looked at how artist started using objects that are not produced by artist and with the purpose of making art. Their works are notable and founding works that open new dimensions to reconsider objects and their meaning. Further, they extended the ways of art making and the language of art. Therefore, it is important to analyze Their intent and methods are analyzed (how?). In other words, looked at the foundation of using non-art objects in the art making. First examples are analyzed and how it evolved over time is explored. Questioned that how they affected successor artist.

Secondly looked at the some examples of contemporary works (or works from recent history). In particular, capture the current approaches on trash. Several methods and genres are explored.

Lastly focused on the inspiring documentary of Agnes Varda, The Gleaners and I which provides a general picture of the subject.

%
%
\section{Root in the Art History}
Using trashed objects in making art can be examined with the entrance of non-art objects into the art. In other words, using objects in different scope beyond their intended purpose or intended meaning and function. Developing artworks not only painting but also using paper and other stuff by pasting (or juxtaposing) them together. (How using non-art objects affected art? Fragments) Using object apart from their proposed meaning is not a new thing, through the ages people used objects and tools for different purposes. However, it is not discussed in the context of fine art and not accepted method for art making.

From the beginning of 20th-century non-art, objects draw the attention of the artist. Development of art in this century evolved through this non-art objects (or these objects play significant role development of art trough time). They bring new dimension (or perception) to consider objects. Usage of these objects dramatically changed the way people look at art. In short, they introduced a new way of making art.

One of the remarkable change in the 20th-century art is the using unconventional materials in the making of paintings and sculptures, rather than the original (pure, fresh) materials produced with the purpose of art. These objects are not normally considered as art, often because they already have a non-art function. Possibly they are part of different disciplines and markets. Their place in the society and people's mind commonly have a different meaning. Beyond them, artists started to use them for their own purposes. Objects gain different meanings and usages with the hands of artists. (It is explained in the previous chapter that objects are not static and absolute. Objects can be used for new purposes even if they lost their meaning and value in the conventional system.)

They questioned the way of traditional (conventional) way of art making in every dimension. Their approach out of the previous methods. They are looking for the beyond of painting. With using different objects they break the pureness of artwork. It is not pure anymore, fragmented. Boundaries are pushed. Using non-art objects in the production of art challenges existing conventions of painting.

Using non-art object can be considered as moving away from the purification of art \todo{ref req}. It is a composition of different fragments and pieces. Rather than the traditional way of painting on canvas they composed their work with the combination of various objects from different context. With this approach works can be examined in multi-dimensional context.

@@

Before the 20th-century, artists started to inquiry the existing methods and themes of previous periods. Modern artists experimented with new ways of seeing and with fresh ideas about the nature of materials and functions of art. Previously paintings are based on linear perspective and realistic style. Themes are often composed of religious and mythical narratives \todo{MOMA ref}. After the second half of the 19th-century, they started to abandon these themes and painting methods. For instance, in the works of impressionist paint brushstrokes can be easily identified, and themes are composed of everyday scenes. With Cezanne, the linear perspective started to question and later reached to its peak point with cubism. Cubism is developed with the collaboration of the Picasso and Braque. Besides, these two artists experimented non-traditional materials in their works for the first time in the scope of fine art. They continuously produced works with using different objects. They support each others art. They showed that an endless potential lay there. They collaborated with each other on the development of cubism and collage.

One of the keys to understanding the importance of Cubism, of Picasso and Braque, is to consider their actions and how unusual they were for the time. When Braque, and then Picasso placed industrially-produced objects ("low" commercial culture) into the realm of fine art ("high" culture) they acted as artistic iconoclasts (icon=image/clast=destroyer). \todo{ref}

@@

Collage originates from the French \textit{coller} is an artistic technique of applying manufactured, printed, or found materials, such as bits of newspaper, fabric, wallpaper, etc., to a panel or canvas, frequently in combination with painting. In about 1912–13 Pablo Picasso and Georges Braque extended this technique, combining fragments of paper, wood, linoleum, and newspapers with oil paint on canvas to form compositions. Pasting paper is not a new technique but using this it in the art making is a revolutionary movement in the  language of art \cite{waldman1992collage}.

Collage was a major turning point in the evolution of Cubism, and therefore a major turning point in the whole evolution of modernist art in this century \cite{greenberg1984collage}.

Similar to the collage, assamblage is produced by the incorporation of three-dimensional elements. In 1961, the exhibition "The Art of Assemblage" was featured at the New York Museum of Modern Art. William C Seitz, the curator of the exhibition, described assemblages as being made up of preformed natural or manufactured materials, objects, or fragments not intended as art materials. \todo{ref.}

@@

One of the earliest and notable examples of collage is Picasso’s \quotes{Still Life with Chair Caning} (1912), in which a piece of oilcloth which is used as tablecloth or self coverings with an imitation chair caning design was pasted onto the painting. This work of Picasso also can be considered as first example of assemblage because of a rope was used to frame the picture. Picasso combined painting with already existing materials. He used real objects to represent instead of painting. As mentioned in the name of the work Picasso give clues of the parts of it. For instance the oval canvas\todo{canvas or oilcloth?} and the rope around it represent the table. Pasted chair caning signifies the chair within the table. On the table folded paper shadow represent newspaper. Further there is a cut out lemon painted onto the table. \paraphrase{The three letters above the scrap of cloth, "JOU," can be understood as both the beginning of the word "JOURNAL," alluding to the customary newspaper lying across the café table, and as the French verb meaning "to play." The new technique of collage allowed new possibilities of playfulness. This was a new way of making art; instead of painting a thing, you could stick whatever it was right onto the canvas.}

@@

Moreover, they questioned the elitism of the art world, which had always dictated the separation of common, everyday experience from the rarefied, contemplative realm of artistic creation. Of equal importance, their work highlighted and separated the role of technical skill from art-making. Braque and Picasso introduced a “fake” element on purpose, not to mislead or fool their audience, but rather to force a discussion of art and craft, of high and low, of unique and mass-produced objects. They ask: “Can this object still be art if I don’t actually render its forms myself, if the quality of the art is no longer directly tied to my technical skills or level of craftsmanship?"

% http://www.nybooks.com/articles/2010/12/09/artist-everything/
% http://www.artchive.com/artchive/S/schwitters.html
After Picasso and Baraque, many artists applied collage and assamblage techniques to their works. Dada artist Kurt Schwitters is one of the productive artists of them. In 1918 created his own form of Dada in Hanover called 'Merz', using rubbish materials such as labels, bus tickets and bits of broken wood in his collages and constructions. The nonsense word 'Merz' is generated from the name of the name of a bank: the Kommerz und Privatbank. Merz was soon to become a kind of brand name for almost all his activities and indeed, from 1922, he even began to refer to himself as Kurt Merz Schwitters or simply Merz.

"The language of Merz now finds common acceptance and today there is scarcely an artist working with materials other than paint who does not refer to Schwitters in some way. In his bold and wide-ranging experiments he can be seen as the grandfather of Pop Art, Happenings, Concept Art, Fluxus, multimedia art and post-modernism." (Gwendolyn Webster)\todo{ref}

% http://www.artchive.com/artchive/S/schwitters.html
Schwitters’ work inspired such post-war pioneers as Jasper Johns, Robert Rauschenberg and Joseph Beuys, and he is now seen as the grandfather of post-1945 art movements, from Pop Art to Fluxus, Conceptual Art to site-specific art, and the forerunner of present day artists such as Thomas Hirschhorn, Gregor Schneider and Rachel Whiteread.\todo{ref}

Merzbau is most important ones. In 1937 after his work had been included in the Degenerate Art Exhibition he fled Germany for Norway. There he created a second Merzbau. In 1940 he found refuge in England where he started a third Merzbau at Ambleside in the Lake District. The first Merzbau was destroyed in the Second World War, the second by fire in 1951 and the third was left unfinished at his death in 1947. It is now preserved in the Hatton Gallery of the University of Newcastle upon Tyne.

@@

% FROM The Ruin and the Ruined in the Work of Kurt Schwitters by Gemma Carroll
\todo{works of Kurt Schwitters} \paraphrase{The German avant-garde was working from ruins literally and metaphorically, and trash was both practically and freely available; to use it was an action that took the ruins of our society, its discarded, to question how meaning is constructed. Schwitters is able to use the ruined, the waste products, as an anthropological exploration of society from both its unpleasant outcomes and its decay. \ldots trash was both practically and freely available; to use it was an action that took the ruins of our society, its discarded, to question how meaning is constructed. As he wrote: ‘It grows more or less according to the principle of a metropolis.’ The Merzbau was itself a city; and just as Marx wrote that it was not the materiality of the object but the social relations that create value, the use of urban detritus in particular, the squalid results of mass-produced human relations, infuses the materiality of Schwitters’ work with an anthropological quality. Material has transformed into information, and ‘how’ has surpassed ‘what’ we see. The grottos in the Merzbau that still reveal this detritus most clearly could not be re-created in Bissegger’s reconstruction because, arguably, they are an exploration of absence, an exploration of ruin.}

% Book: Recycled, Re-Seen: Folk Art from the Global Scrap Heap
\paraphrase{Like collage in art or quotation in literature, the recycled object carries a kind of "memory" of its prior existence. Recycling always implies a stance toward time --- between the past and the present --- and often a perspective on cultures --- one's own and others. (Jacknis 1992)} \todo{ref.}

Recycled object contains within it a reference to two or more distinct times. Whatever their ultimate design or destination, these recycled artifacts are, by definition, "impure", "inauthentic" products of past and present, here and there, "us" and "them". \todo{ref}

% Book: Recycled, Re-Seen: Folk Art from the Global Scrap Heap 
It is claimed that \paraphrase{the end result is a category of hybrid objects that bear the mark of at least two distinct domains, each with its won material, meaning, makers, and users (p.10).} \paraphrase{Whatever their ultimate function, each of these objects contains within itself a visual, material, and conceptual reference to multiple technologies, histories and temporalities.}

Collages and assemblages can be composed of all types of objects such as trash and found object. The notion of found object is researched by the many scholars \todo{ref.}, but usage in the art has gained new dimensions with the Duchamp's concept of readymade. Industrially manufactured or anything found  anymore can found a place in the art. The most famous example is Fountain (1917) (discussed in the previous chapter), a standard urinal purchased from a hardware store and displayed on a pedestal, resting on its side. Before that first one is Bicycle Wheel. He mounted the bicycle wheel upside down onto a stool. 
% http://www.moma.org/learn/moma_learning/marcel-duchamp-bicycle-wheel-new-york-1951-third-version-after-lost-original-of-1913
When Bicycle Wheel was first displayed, Duchamp encouraged viewers to spin the wheel. (Participation of the audience) Although he claimed to select objects for his readymades without regard to beauty, he said, “To see that wheel turning was very soothing, very comforting\ldots I enjoyed looking at it, just as I enjoy looking at the flames dancing in a fireplace.”

@@

%
%
\summary{Found Object} 
Found object originates from the French \textit{objet trouvé}, describing art created from undisguised, but often modified, objects or products that are not normally considered art, often because they already have a non-art function.

@@

\paraphrase{Gascoygne (1936), writing about the artist’s use of “the strange medley of materials” (p. 169), referred to as objets trouve in Surrealist art, suggesting that the artist “discovers a hidden symbolic significance in the [found] object which is preserved when the object is ‘framed’ as art” (p. 170). The finder discovers an unrealised significance in the object. A new boundary is formed around the object by the finder through removing it from its found environment and placing it in a new one, thus empowering the finder in the role of creating a new reality for the object. He argued that the found object, before it is found, approximates a zero value aesthetically; the zero value increases for the finder---beholder on discovery of the object and increases further if the object is placed in another context. \ldots the meaning of material objects was derived from their symbolic relation to another (e.g., person, time, place, experience) rather than through their physical attributes (Causey, 2003; Csikszentmihalyi \& Rochberg-Halton, 1981).} \todo{parap. ref.}

@@

%
%
\section{Examples from contemporary artists}
In the light of discussions at previous section, some of the works (or current strategies and approaches ) of artist are analyzed. First one is related with found objects.

@@

Photobooths are machines that place crowded places and with affordable cost people can get their photos. It is a pheonema that attract attention of artist and researchers such as Martin Parr and Gerry Badger who published history of them as a two volume.

In the French film Amélie\footnote{Original title: Le fabuleux destin d'Amélie Poulain, in english: The Fabulous Destiny of Amélie Poulain} directed by Jean-Pierre Jeunet in 2001, young male character collects photos that are torn up and thrown away by their owner near photo booths spread through train and metro stations of Paris. He collects them and glue them together in a photo album. In the film photo booths and found photos are central catalytic. Amélie falls in love with him when she realize him searching for found photos. After his photo album passed to Amélie, she looks every page of photo album again and again. Through the images of unknown people, she establishes narratives that point out the significance of these images even if they are thrown away. The important part of the film it generates images of a man who scanvage the photos of another people. There are a lot of artist who collects discarded items but their process is not well documented. Therefore we are not aware of the artist process.

@@

Prior to film Amélie, British artist Dick Jewell published a book in title of "Found Photos" in 1977. It is a collection of photobooth images that had been thrown away or torn up by the people in the photos. Ian Walker comments on that it is "always a joy to look through." Images capture diverse range of activities and expressions of people using photo booth machine. They are scavenged by the artist from the photo booths of London. The reasons of why photos are rejected vary including technical problems like flash or posing time. On the other hand, some of them are very good condition, and it is a mystery that why they are rejected. Walker explains that the \todo{block qoute} "gap between our exterior perception of the image and what must have been the internal dissatisfaction of the original subject results in some of the effective moments in Found Photos, where the images are not only bizarre and wonderfully funny (as many of them are) but also profoundly moving." Walker argues that even if photo booths are mechanic and simple application of photography while capturing everyday life, leave behind such ambiguous and mysteries that some people trace them. "Dick Jewell’s approach with Found Photos was less interventionist, presenting the images themselves in the spirit of either an anthropologist showing his finds or a conceptual artist claiming bits of the real world as ‘readymades’." His mode of discovery includes climbing to look at the top of the box and looking for the nearest waste bin. Badger describes Found Photos as ‘both a conceptual artwork and a sociological footnote’.

@@

\begin{wrapfigure}{r}{0.4\textwidth}
  \begin{center}
    \includegraphics[width=0.37\textwidth]{graphics/vik-muniz-picturesofgarbage0.jpg}
  \end{center}
  \caption{Vic Muniz, Pictures of Garbage}
  \label{fig:VicMuniz_PicturesOfGarbage}
\end{wrapfigure}

Not only monumental portraits extracted from the dump but also an orchestra is grounded up from the dump of Paraguay pioneered by Favio Chávez and Nicolás Gómez. Favio Chávez is a music teacher, and Nicolás Gómez is a garbage picker. They live in Cateura at Paraguay. It is a one of the poorest villages where people live among the sea of garbage. "A violin is worth more than a house here," says Favio Chavez in documentary film. The orchestra's director and founder. Favio Chavez director and founder of the orchestra whose instruments such as violins, flutes and cellos are built from collected trash by Nicolás Gómez. Players of the orchestra are compiled from the children who are hopeless for their future. In the midst of such an existence, these musicians have created something both special and truly awe-inspiring. They are given birth to an orchestra that plays the masterpieces of classical music composed by Mozart, Beethoven, and Vivaldi. Trash turned to treasure and totally ironic that a signifiers of high culture like classical music can be even played with imperfect instruments. For them and also, most of the people it is not affordable to buy an instrument and get the education of classical music. (To enjoy and play such music requires economic wealth and it is not possible for everyone.) What they create is actually not expected from them, and their success is not limited with their location. They have given concerts around the world.

@@

They build their musical instruments from what is available around. This project can be seen as a bricolage practice. Dictionary meaning\todo{which dictionary}; something constructed using whatever was available at the time. \paraphrase{Claude Levi-Strauss notes that the \textit{bricoleur} works not from the principle of making things only if natural resources are available but makes things according to those things at hand, making do with what is available. It is an expression that, like the natural cycles of the Earth, attempts to make something new from something old. \cite{levi1966savage}} The \textit{bricoleur} is \ldots someone who works with his [or her] hands and uses devious means compared to those of a craftsman \ldots \cite{levi1966savage}

@@

In the scope of painting different than the works in the previous chapter (paintings on tea bags and coffee cups) historical potraits of notable people are painted onto the flattened aluminum drink cans which are disposable objects. The people in the figure X is General Jefferson C. Davis who was an officer at United States Army in 19th-century. Selected persons can be viewed as a member of the higher or ruling class of society. Therefore Kim Alsbrooks, painter of these works, might thought that to make trash is more valuable, their face will be good. Or the vice versa. She may use trash to provocate social elite of that time.

\begin{wrapfigure}{r}{0.4\textwidth}
  \begin{center}
    \includegraphics[width=0.37\textwidth]{graphics/Alsbrooks.jpg}
  \end{center}
  \caption{Kim Alsbrooks, The White Trash Series}
  \label{fig:Alsbrooks}
\end{wrapfigure}

>>>>

\chapter{TRASH (IN) ART}

% Epigraph
\begin{singlespace}
\epigraph{One day, in a rubbish heap, I found an old bicycle seat lying beside a rusted handlebar, and my mind instantly linked them together. I assembled these two objects, which everyone then recognized as a bull’s head. The metamorphosis was accomplished, and I wish another metamorphosis would occur in the reverse sense. If my bull’s head were thrown in a junk heap, perhaps one day some boy would say, \quotes{Here’s something that would make a good handlebar for my bicycle!}}{\hfill ---Pablo Picasso, \textit{Trashformations}, 1998}
\end{singlespace}

In this chapter, I question the purpose and meaning of using trash as a medium in the artworks. The reasons and motivations of artists who use trash will be elaborated. The artists who use discarded items in their work will be analyzed in the light of arguments discussed in the previous chapter. Here, the main question is why artists select to use society's debris rather than fresh items. Hence, my research will focus on the place of trash in the world of art by looking at different methods and approaches. The scope of this research is Started from masterpieces and founding works, and continued with recent examples of artworks.

@@

Firstly I look at how artist started using objects that are not produced by herself/himself and with the purpose of making art. Early works are notable and founding ones that open new dimensions to reconsider objects and their meaning. Further, they extended the ways of art making and the language of art. Therefore, it is important and necessarry to analyze them. In other words, I first look at the foundation of using non-art objects in the art making. Early examples are analyzed and how they evolved over time is explored. I try to find out how they affected successor artist.

Secondly looked at some examples of contemporary works (or works from recent history). In particular, I try to capture the current approaches on trash. Several methods and genres are explored.  Lastly focused on the inspiring documentary of Agnes Varda, The Gleaners and I which provides a general picture of the subject.

%
%
\section{Root in the Art History}
Using trashed objects in making art can be examined with the use of non-art objects in modern art. In other words, some artist started using objects in different scope beyond their intended purpose or intended meaning and function in their modern artworks. They develop their artworks by using paper and other stuff by pasting (or juxtaposing) them together.

%Indeed, using object apart from their proposed meaning is not a new thing, through the ages people used objects and tools for different purposes. However, it is not discussed in the context of fine art and not accepted method for art making.

From the beginning of 20th-century non-art objects draw the attention of the artist. Development of art in this century evolved through this non-art objects (or these objects play significant role development of art trough time). Usage of these objects dramatically changed the way how people look at art. In short, they introduced a new way of making art.

One of the remarkable change in the 20th-century art is the use of unconventional materials in the making of paintings and sculptures, rather than the conventional materials produced with the purpose of art. These objects are not normally considered as art often because they already have a non-art function. Their place in the society and people's mind commonly have a different meaning. Beyond that, artists started to use them for their own purposes. Thus, objects gained different meanings and usages with the hands of artists. (It is explained in the previous chapter that objects are not static and absolute. Objects can be used for new purposes even if they lost their meaning and value in the conventional system.)

Modern artist questioned the way of traditional (conventional) way of art making in every dimension. Their approach out of the previous methods. They are looking for the beyond of painting. With using different objects they break the pureness of artwork. It is not pure anymore, fragmented. Boundaries are pushed. Using non-art objects in the production of art challenges existing conventions of painting.

Using non-art object can be considered as moving away from the purification of art \todo{ref req}. It is a composition of different fragments and pieces. Rather than the traditional way of painting on canvas modern artists composed their work with the combination of various objects from different context. With this approach works can be examined in multi-dimensional context.

@@

In the 20th-century, modern artists experimented with new ways of seeing and with fresh ideas about the nature of materials and functions of art. Previously paintings are based on linear perspective and realistic style. Themes are often composed of religious and mythical narratives. After the second half of the 19th-century, they started to abandon these themes and painting methods. For instance, in the works of impressionist paint brushstrokes can be easily identified. In impressionist paintings, themes are composed of everyday scenes. With Cezanne, the linear perspective is questioned and later this questioning reached to its peak point with cubism. Cubism is developed with the collaboration of Picasso and Braque. Besides, these two artists experimented non-traditional materials in their works for the first time in the scope of fine art. They continuously produced works with using different objects. They support each others art. They showed that an endless potential lay there. They collaborated with each other on the development of cubism and collage.

One of the keys to understanding the importance of Cubism of Picasso and Braque, is to consider their actions and how unusual they were for the time. When Braque, and then Picasso placed industrially-produced objects into the realm of fine art they acted as artistic iconoclasts of their age. 
%\todo{ref}

@@

Collage originates from the French \textit{coller} is an artistic technique of applying manufactured, printed, or found materials, such as bits of newspaper, fabric, wallpaper, etc., to a panel or canvas, frequently in combination with painting. In about 1912–13 Pablo Picasso and Georges Braque extended this technique, combining fragments of paper, wood, linoleum, and newspapers with oil paint on canvas to form compositions. Pasting paper is not a new technique but using this it in the art making is a revolutionary movement in the  language of art \citep{waldman1992collage}. Collage was a major turning point in the evolution of Cubism, and therefore a major turning point in the whole evolution of modernist art in this century \citep{greenberg1984collage}.

Similar to the collage, assamblage is produced by the incorporation of three-dimensional elements. In 1961, the exhibition "The Art of Assemblage" was featured at the New York Museum of Modern Art. William C Seitz, the curator of the exhibition, described assemblages as being made up of preformed natural or manufactured materials, objects, or fragments not intended as art materials. 
%\todo{ref.}

@@

One of the earliest and notable examples of collage is Picasso’s \quotes{Still Life with Chair Caning} (1912), in which a piece of oilcloth which is used as tablecloth or self coverings with an imitation chair caning design was pasted onto the painting. This work of Picasso also can be considered as first example of assemblage because of a rope that was used to frame the picture. Picasso combined painting with already existing materials. As mentioned in the name of the work, Picasso give clues of the parts of it. For instance the oval canvas and the rope around it represent the table. Pasted chair caning signifies the chair within the table. On the table folded paper shadow represent newspaper. Further there is a cut out lemon painted onto the table. \paraphrase{The three letters above the scrap of cloth, "JOU," can be understood as both the beginning of the word "JOURNAL," alluding to the customary newspaper lying across the café table, and as the French verb meaning "to play." The new technique of collage allowed new possibilities of playfulness. This was a new way of making art; instead of painting a thing, you could stick whatever it was right onto the canvas.}

@@

Moreover, Braque and Picasso questioned the elitism of the art world, which had always dictated the separation of common, everyday experience from the  contemplative realm of artistic creation. Of equal importance, their work highlighted and separated the role of technical skill from art-making. Braque and Picasso introduced a “fake” element on purpose, not to mislead or fool their audience, but rather to force a discussion of art and craft, of unique and mass-produced objects. They ask: “Can this object still be art if I don’t actually render its forms myself, if the quality of the art is no longer directly tied to my technical skills or level of craftsmanship?"

% http://www.nybooks.com/articles/2010/12/09/artist-everything/
% http://www.artchive.com/artchive/S/schwitters.html
After Picasso and Baraque, many artists applied collage and assamblage techniques to their works. Dada artist Kurt Schwitters is one of the productive artists amoung them. In 1918 created his own form of Dada in Hanover called 'Merz', using rubbish materials such as labels, bus tickets and bits of broken wood in his collages and constructions. The nonsense word 'Merz' is generated from the name of the name of a bank: the Kommerz und Privatbank. Merz was soon to become a kind of brand name for almost all his activities and indeed, from 1922, he even began to refer to himself as Kurt Merz Schwitters or simply Merz.

\begin{singlespace}
\begin{quote}
The language of Merz now finds common acceptance and today there is scarcely an artist working with materials other than paint who does not refer to Schwitters in some way. In his bold and wide-ranging experiments he can be seen as the grandfather of Pop Art, Happenings, Concept Art, Fluxus, multimedia art and post-modernism...

Schwitters’ work inspired such post-war pioneers as Jasper Johns, Robert Rauschenberg and Joseph Beuys, and he is now seen as the grandfather of post-1945 art movements, from Pop Art to Fluxus, Conceptual Art to site-specific art, and the forerunner of present day artists such as Thomas Hirschhorn, Gregor Schneider and Rachel Whiteread. \citep{webster2011kurt}
\end{quote}
\end{singlespace}

In 1937 after Merzbau had been included in the Degenerate Art Exhibition he fled from Germany to Norway. There he created a second Merzbau. In 1940 he found refuge in England where he started a third Merzbau at Ambleside in the Lake District. The first Merzbau was destroyed in the Second World War, the second by fire in 1951 and the third was left unfinished at his death in 1947. It is now preserved in the Hatton Gallery of the University of Newcastle upon Tyne.

@@

% FROM The Ruin and the Ruined in the Work of Kurt Schwitters by Gemma Carroll
\begin{singlespace}
\begin{quote}
The German avant-garde was working from ruins literally and metaphorically, and trash was both practically and freely available; to use it was an action that took the ruins of our society, its discarded, to question how meaning is constructed. Schwitters is able to use the ruined, the waste products, as an anthropological exploration of society from both its unpleasant outcomes and its decay. \ldots trash was both practically and freely available; to use it was an action that took the ruins of our society, its discarded, to question how meaning is constructed. As he wrote: ‘It grows more or less according to the principle of a metropolis.’ The Merzbau was itself a city; and just as Marx wrote that it was not the materiality of the object but the social relations that create value, the use of urban detritus in particular, the squalid results of mass-produced human relations, infuses the materiality of Schwitters’ work with an anthropological quality. Material has transformed into information, and ‘how’ has surpassed ‘what’ we see. The grottos in the Merzbau that still reveal this detritus most clearly could not be re-created in Bissegger’s reconstruction because, arguably, they are an exploration of absence, an exploration of ruin. \citep{carroll2011ruin}
\end{quote}
\end{singlespace}

% Book: Recycled, Re-Seen: Folk Art from the Global Scrap Heap
Jacknis claims that \quotes{like collage in art or quotation in literature, the recycled object carries a kind of "memory" of its prior existence. Recycling always implies a stance toward time --- between the past and the present --- and often a perspective on cultures --- one's own and others} \citep[as cited in][]{cerny1996recycled}

Recycled object contains within it a reference to two or more distinct times. Whatever their ultimate design or destination, these recycled artifacts are, by definition, "impure", "inauthentic" products of past and present, here and there, "us" and "them". 
%\todo{ref}

% Book: Recycled, Re-Seen: Folk Art from the Global Scrap Heap 
It is claimed that the end result is a category of hybrid objects that bear the mark of at least two distinct domains, each with its won material, meaning, makers, and users \citep{cerny1996recycled} Whatever their ultimate function, each of these objects contains within itself a visual, material, and conceptual reference to multiple technologies, histories and temporalities.

Collages and assemblages can be composed of all types of objects such as trash and found object. The notion of found object is researched by many scholars 
%\todo{ref.}
, but usage in the art has gained new dimensions with the Duchamp's concept of readymade. Industrially manufactured or anything found  anymore can found a place in the art. The most famous example is Fountain (1917) (discussed in the previous chapter), a standard urinal purchased from a hardware store and displayed on a pedestal, resting on its side. Before that first one is Bicycle Wheel. He mounted the bicycle wheel upside down onto a stool. 
% http://www.moma.org/learn/moma_learning/marcel-duchamp-bicycle-wheel-new-york-1951-third-version-after-lost-original-of-1913
When Bicycle Wheel was first displayed, Duchamp encouraged viewers to spin the wheel. Although he claimed to select objects for his readymades without regard to beauty, he said, “To see that wheel turning was very soothing, very comforting\ldots I enjoyed looking at it, just as I enjoy looking at the flames dancing in a fireplace.”

@@

%
%
%\summary{Found Object} 
Found object originates from the French \textit{objet trouvé}, describing art created from undisguised, but often modified, objects or products that are not normally considered art, often because they already have a non-art function.

@@

Gascoygne (1936), writing about the artist’s use of “the strange medley of materials” (p. 169), referred to as objets trouve in Surrealist art, suggesting that the artist “discovers a hidden symbolic significance in the [found] object which is preserved when the object is ‘framed’ as art” (p. 170). The finder discovers an unrealised significance in the object. A new boundary is formed around the object by the finder through removing it from its found environment and placing it in a new one, thus empowering the finder in the role of creating a new reality for the object. He argued that the found object, before it is found, approximates a zero value aesthetically; the zero value increases for the finder---beholder on discovery of the object and increases further if the object is placed in another context. \ldots the meaning of material objects was derived from their symbolic relation to another (e.g., person, time, place, experience) rather than through their physical attributes (Causey, 2003; Csikszentmihalyi \& Rochberg-Halton, 1981). 
%\todo{parap. ref.}

@@

%
%
\section{Examples from contemporary artists}
In the light of discussions at previous section, some of the main strategies and approaches of artist are analyzed. First one is related with found objects. In this section focus will be shifted to the examples from contemporary artists.

@@

Photobooths are machines that place crowded places and with affordable cost people can get their photos. It is a pheonema that attract attention of artist and researchers such as Martin Parr and Gerry Badger who published history of them as two volume.

In the French film Amélie\footnote{Original title: Le fabuleux destin d'Amélie Poulain, in english: The Fabulous Destiny of Amélie Poulain} directed by Jean-Pierre Jeunet in 2001, young male character collects photos that are torn up and thrown away by their owner near photo booths spread through train and metro stations of Paris. He collects them and glue them together in a photo album. In the film photo booths and found photos are centrally catalytic. Amélie falls in love with him when she realize him searching for found photos. After his photo album passed to Amélie, she looks every page of photo album again and again. Through the images of unknown people, she establishes narratives that point out the significance of these images even if they are thrown away. The important part of the film, images of a man are generated who scanvage the photos of another people. There are a lot of artist who collects discarded items but their process is not well documented. Therefore we are not aware of the artistic process.

@@

Prior to film Amélie, British artist Dick Jewell published a book in title of "Found Photos" in 1977. It is a collection of photobooth images that had been thrown away or torn up by the people in the photos. \cite{walker2010dick} comments on that it is "always a joy to look through." Images capture diverse range of activities and expressions of people using photo booth machine. They are scavenged by the artist from the photo booths of London. The reasons of why photos are rejected vary including technical problems like flash or posing time. On the other hand, some of them are very good condition, and it is a mystery that why they are rejected. Walker explains that the

\begin{singlespace}
\begin{quote}
gap between our exterior perception of the image and what must have been the internal dissatisfaction of the original subject results in some of the effective moments in Found Photos, where the images are not only bizarre and wonderfully funny (as many of them are) but also profoundly moving.
\end{quote}
\end{singlespace}

Walker argues that even if photo booths are mechanic and simple application of photography while capturing everyday life, leave behind such ambiguous and mysteries that some people trace them. "Dick Jewell’s approach with Found Photos was less interventionist, presenting the images themselves in the spirit of either an anthropologist showing his finds or a conceptual artist claiming bits of the real world as ‘readymades’." His mode of discovery includes climbing to look at the top of the box and looking for the nearest waste bin. Badger describes Found Photos as ‘both a conceptual artwork and a sociological footnote’.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[height=8cm]{graphics/vik-muniz-picturesofgarbage0.jpg}
  \caption{Vic Muniz, Pictures of Garbage}
  \label{fig:VicMuniz_PicturesOfGarbage}
\end{figure}

Not only monumental portraits extracted from the dump but also an orchestra is grounded up from the dump of Paraguay pioneered by Favio Chávez and Nicolás Gómez. Favio Chávez is a music teacher, and Nicolás Gómez is a garbage picker. They live in Cateura at Paraguay. It is a one of the poorest villages where people live among the sea of garbage. "A violin is worth more than a house here," says Favio Chavez in documentary film. The orchestra's director and founder. Favio Chavez director and founder of the orchestra whose instruments such as violins, flutes and cellos are built from collected trash by Nicolás Gómez. Players of the orchestra are compiled from the children who are hopeless for their future. In the midst of such an existence, these musicians have created something both special and truly awe-inspiring. They are given birth to an orchestra that plays the masterpieces of classical music composed by Mozart, Beethoven, and Vivaldi. Trash turned to treasure and totally ironic that a signifiers of high culture like classical music can be even played with imperfect instruments. For them and also, most of the people it is not affordable to buy an instrument and get the education of classical music. What they create is actually not expected from them, and their success is not limited with their location. They have given concerts around the world.

@@

They build their musical instruments from what is available around. This project can be seen as a bricolage practice. Dictionary meaning; something constructed using whatever was available at the time. \paraphrase{Claude Levi-Strauss notes that the \textit{bricoleur} works not from the principle of making things only if natural resources are available but makes things according to those things at hand, making do with what is available. It is an expression that, like the natural cycles of the Earth, attempts to make something new from something old. \cite{levi1966savage}} The \textit{bricoleur} is \ldots someone who works with his [or her] hands and uses devious means compared to those of a craftsman \citep{levi1966savage}.

@@

In the scope of painting different than the works in the previous chapter (paintings on tea bags and coffee cups) historical potraits of notable people are painted onto the flattened aluminum drink cans which are disposable objects. The people in the figure X is General Jefferson C. Davis who was an officer at United States Army in 19th-century. Selected persons can be viewed as a member of the higher or ruling class of society. Therefore Kim Alsbrooks, painter of these works, might have thought that to make trash is more valuable, their face will be good. Or vice versa. She probably use trash to provocate social elite of that time.

\begin{figure}[h!]
  \centering
  \includegraphics[height=8cm]{graphics/Alsbrooks.jpg}
  \caption{Kim Alsbrooks, The White Trash Series}
  \label{fig:Alsbrooks}
\end{figure}

Not all the time and every artist turn trash into different objects, sometimes they only trash them. Michael Landy's Art Bin is one of them. It is a room size transparent bin placed inside of the art gallery. He invites people to bring their own artworks of failed attempts. Some of the notable contemporary artists such as Damien Hirst, Gillian Wearing, Tracey Emin and Mark Titchner participated in by throwing their sculptures, paintings, and prints into the Art Bin. People publicly toss their work from a high platform, and all of them can be seen outside of the transparent walls. He listens to the story and the failure of the artworks from their creator, and it is not allowed to throw others work. At one side he criticizes the idea of not aware of that artist also have lots of failures. People only can see the final work in the galleries and museums. In this work he break-ups this notion that encouraging the artist reveals their failures to the public. Like his words, it is “a monument to creative failure”. On the other side, there is a provocative approach to the art and art objects by saying that "Nothing is too good for the art bin." There is no limitation of throwing away even if art.
